# Losistic Regression (Excercise 3)
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
The data is about student acheivement in secondary education of two Portugues schools. Two datasets are available for the performance in two subjects: Math and Portuguese language.  

P. Cortez and A. Silva. Using Data Mining to Predict Secondary School Student Performance. In A. Brito and J. Teixeira Eds., Proceedings of 5th FUture BUsiness TEChnology Conference (FUBUTEC 2008) pp. 5-12, Porto, Portugal, April, 2008, EUROSIS, ISBN 978-9077381-39-7.[link]http://www3.dsi.uminho.pt/pcortez/student.pdf

### 1. Retreiving data 
the data is about student acheivement in secondary education of two Portugues schools according to the student grades, demographic and school-related features. Two datasets are available for the performance in two subjects: Math and Portuguese language. the datasets consists of 35 variables and 382 observations.I firstly beging with installing some packages helps analysis. And thus, we can take a glimpse at the datasets and what variables are available through column names on the datasets.
```{r}
library(dplyr);library(ggplot2);library(tidyr);library(boot)
alc<-read.csv("Z:\\IODS-project\\data\\alc.csv")
glimpse(alc)
gather(alc)%>% glimpse
gather(alc)%>%ggplot(aes(value))+facet_wrap("key", scales="free")+geom_bar()
```

### 2. Hypothesis  
In order to study the relationship between high and low alcohol consumptions and variables, i am going to work folloing to hypothesis;

1. **Gender** - Males consume more alcohol than females.
2. **G1** - First period grade decreases with alcohol consumption.
3. **G2** - Second period grade decreases with alcohol consumption.
4. **G3** - Third period grade decreases with alcohol consumption.

I will take a look at each relationship followring to the hypothesis one by one
p<-ggpairs(alc, mapping = aes(col=sex, alpha = 0.3), lower = list(combo = wrap("facethist", bins = 20)))
p

#### Hypothesis 1 - **Gender**

```{r}

g1 <- ggplot(data = alc, aes(x = alc_use, fill = sex))
g1 + geom_bar()+facet_wrap("sex") + ggtitle("alohol consumption by sex")
g11<-ggplot(data=alc, aes(x=high_use, fill=sex))
g11+geom_bar()+facet_wrap("sex") + ggtitle("high and low alcohol consumption by sex")

```


#### Hypothesis 2 - **G1**
```{r}
g2 <- ggplot(data=alc, aes(x=high_use, y= G1, col=sex))
g2+geom_boxplot()+ggtitle("G1 by alcohol consumption and sex")
```

#### Hypothesis 3 - **G2**
```{r}
g3<-ggplot(alc, aes(x=high_use, y=G2, col=sex))
g3+geom_boxplot()+ggtitle("G2 by acohol consumption and sex")
```

#### Hypothesis 4 - **Absences**
```{r}
g4<-ggplot(alc, aes(x=high_use, y=G3, col=sex))
g4+geom_boxplot()+ggtitle("G3 by alcohol consumption and sex")
```

#### Interim result
As a result of exploration of relationship with 4 variable to the high and low alcohol relationship, it seems that femailes consume more alcohol than males do. However, it is presumed that the more males consum alcohol, the less grades males achieve. 


### 3.Lostistic regression
From now, in order to interpret relationship between variables presumably relating to alcohol consumptions and alcohol consumption, we will deal with logistaic regression.

```{r}
m<-glm(high_use~sex+G1+G2+G3, data=alc, family="binomial")
summary(m)
coef(m)

OR<-coef(m)%>%exp
OR
CI<-exp(confint(m))
CI
cbind(OR, CI)
```

The model indicates taht males are more significantly related to high alcohol consumptions.However, the other 3 variables are not significantly related to high alcohol consumptions.
In odds rates, the odds rates for G2 and G3 are more than 1. 

#### Prediction power 

```{r}
probabilities<-predict(m, type="response")
alc<-mutate(alc,probability=probabilities)
alc<-mutate(alc,prediction=probability>0.5)
select(alc,G2,G3, high_use, probability, prediction)%>% tail(10)
```

#### 2x2 cross-tabulation and graphic representation
```{r}
g<-ggplot(alc,(aes(x=probability, y=high_use, col=prediction)))
g+geom_point()
table(high_use=alc$high_use, prediction=alc$prediction)%>%prop.table()%>%addmargins()
```
#### Exploring the training error
```{r}
loss_func<-function(class, prob){
  n_wrong <-abs(class - prob) > 0.5
  mean(n_wrong)
}
loss_func(class = alc$high_use, prob = alc$probability)
```

Through exploring the training error, the incorrection of prediction is almost 50%.

#### 10-fold cross-validation (Bonus task)
```{r}
cv<-cv.glm(data=alc, cost=loss_func, glmfit = m, K=10)
cv$delta[1]
```


#### Eventually, the prediction error with the 4 variables is **40%**.